---
title: "STAT 240 - Assignment 3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(RSQLite)

dbcon = dbConnect(SQLite(), dbname="lab03.sqlite")
```

## Problem 1
```{r, echo=TRUE}
query1 = "SELECT COUNT(DISTINCT year) AS DistinctYears FROM WinterO"
dbGetQuery(dbcon, query1)
```

## Problem 2
```{r, echo=FALSE}
query2 = "SELECT DISTINCT Height_m FROM Pokem ORDER BY Height_m DESC"
dbGetQuery(dbcon, query2)
```

## Problem 3
# a
```{r, echo=TRUE, warning=FALSE}
query3 = "SELECT * FROM CA"
system.time(for(i in 1:10000) {dbSendQuery(dbcon, query3)} )
```
# b
```{r, echo=TRUE, warning=FALSE}
query_out3 = dbSendQuery(dbcon, query3)
system.time(for(i in 1:10000) {dbFetch(query_out3, 1)} )
dbClearResult(query_out3)
```
# c
'dbFetch' method was much faster than 'dbSendQuery'. And that's because 'dbfetch' gets a single row at a time from the query result whereas 'dbSendQuery' sends the query each time.

## Problem 4
The LIKE operator matches the given pattern with strings in the rows of the query result. Special characters such as % , _ , and \\\ along with other arbitrary characters are used to specify the requirements. For example, % character stands for 0 or more arbitrary characters, _ character stands for exactly 1 arbitrary characters, and \\\ character is used to escape the special characters % and _ .

# Selects all rows from table CA where column Geographic_names /postal code are 3 letters long and starts with 'V4' and then one single character.
```{r, echo=TRUE}
query41 = "SELECT * FROM CA WHERE Geographic_name LIKE 'V4_'"
query_out41 = dbSendQuery(dbcon, query41)
dbFetch(query_out41, 5)
dbClearResult(query_out41)
```
# Selects all rows from table CA where column Region has brackets '()' in the  Region / city names.
```{r, echo=TRUE}
query42 = "SELECT * FROM CA WHERE Region LIKE '%(%)'"
query_out42 = dbSendQuery(dbcon, query42)
dbFetch(query_out42, 5)
dbClearResult(query_out42)
```
